---
title: "Credit Scoring Tools"
author: "Adrien Ehrhardt"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Credit Scoring Tools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# What is Credit Scoring?

This package has been developed as part of a CIFRE PhD, a special PhD contract in France which is for the most part financed by a company. This company subsequently gets to choose which subject(s) are tackled.



## Context


## Formulation


## Conceptual problems of current approaches to Credit Scoring


## Problems tackled in this package


# Reject Inference

## Context

## Theoretical findings

## Methods


# Discretization

## Context

## Mathematical reinterpretation

## Existing methods

## Proposed method




## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
